---
title: "R 101"
output: html_document
date: "2023-05-13"
---

#### 1.1 包
```{r}
#install.packages("tidyverse") 下载package
#library("tidyverse") 用package
```
#### 1.2 导入dataset
```{r}
#dataset = read.csv("path") 把data读进r（file -> import dataset -> from text (base))
```
#### 1.3 rmd路径
```{r}
getwd() #拿到此rmd的路径
```

#### 2.1 dataset的大小
```{r}
dim(iris) #看dataset有几行 几列 (rows and columns)
```
#### 2.2 variables的名字
```{r}
names(iris) #拿到dataset的variables的名字 (column names)
#names(iris) = c('x1', 'x2', 'x3', 'x4', 'x5') 改这些名字
```
#### 2.3 选几行来看
```{r}
head(iris)  #看前6行data（6是r默认的，可以自己改 by adding n = 'x'）
tail(iris, n=3)  #看最后3行data
iris[5,1] #拿到第五行第一个data
```
#### 2.4 dataset大概有些啥
```{r}
str(iris) #看dataset有多少个object，多少个variable，都是什么类型 (eg. chr, num, int)
```
#### 2.5 提取一行数据出来
```{r}
var1 = iris$Sepal.Length #把var1这个variable从dataset提取出来，并起名叫var1
head(var1) #看前六个data
class(var1) #看这个variable是什么类型
length(var1) #看有多少个data
sum(var1) #所有var1的data的和，如果得到NA说明有missing value
```
#### 2.6 排序
```{r}
sort(var1) #从小到大把dataset排序
sort(var1, decreasing = TRUE) #从大到小
sum(sort(var1)[1:5]) #拿到最小的五个data的和
```
#### 2.7 类型变换
```{r}
as.character(iris$Sepal.Length[1:5]) #把numerical变成categorical，改了前五行要不然太多数字看着头疼
as.numeric(iris$Sepal.Length[1:5]) #把categorical变成numerical
```

#### 3.1 basic r 画图
```{r}
table(iris$Sepal.Length) #每个category在数据中出现的次数
barplot(iris$Petal.Length[1:20]) #前20个数据画成bar plot，全画的话会密集恐惧症
hist(iris$Petal.Length) #画histogram
hist(iris$Petal.Length, freq = FALSE) #用probability densities画histogram
boxplot(iris$Sepal.Length, horizontal = TRUE) #把box plot横过来
plot(iris$Petal.Width, iris$Petal.Length) #scatter plot, 前面x后面y
```


#### 3.2 ggplot 画bar plot
```{r}
library(ggplot2)
#bar plot
ggplot(iris, aes(x=factor(Sepal.Width), fill = Species)) + geom_bar() + #Sepal.Width是numerical所以要factor
  ggtitle("Sepal Width by Species") + xlab("Sepal Width")
```

#### 3.3 ggplot 画 histogram
```{r}
ggplot(iris, aes(x = Sepal.Width)) + geom_histogram(bins = 15)
```

#### 3.4 ggplot 画 comparative box plot
```{r}
ggplot(iris, aes(x = Species, y = Petal.Width, fill = Species)) + geom_boxplot() + 
  theme(legend.position = 'none')
```

#### 4.1 把r当计算器 -> Numerical summaries
```{r}
mean(iris$Sepal.Length)
median(iris$Sepal.Length)
sd(iris$Sepal.Length)

library(rafalib)
popsd(iris$Sepal.Length)

var(iris$Sepal.Length)
IQR(iris$Sepal.Length)
```
#### 4.2 Normal model
```{r}
#已知z-score是0.8，算normal curve下0.8左边的面积：
pnorm(0.8) #后面不写mean和sd的话默认standard normal curve (mean = 0, sd = 1)
pnorm(0.8, lower.tail = FALSE) #算0.8右边的面积，也可以
1-pnorm(0.8)
pnorm(0.6) - pnorm(0.1) #算这两个值中间的面积
pnorm(180, mean = 177, sd = 5) #自定义mean和sd
pnorm(180, 177, 5) #也可以直接写数字

qnorm(0.3, 177, 5) #已知面积，算z-score（也可以理解为 百分之三十的数据小于多少）

curve(dnorm(x), from = -3*1, to = 3*1, 
    xlab = "x", ylab = "Density", main = "Normal Distribution")

```


#### 5.1 Linear model
```{r}
#两个numerical data -> scatter plot
ggplot(iris,aes(x = Petal.Width, y = Petal.Length, colour = Species)) + geom_jitter() + 
  labs(x = "Petal Width", y = "Petal Length") + ggtitle("Petal Width vs Petal Length") + 
  geom_smooth(method = "lm", col = "skyblue", se = FALSE) 
linear_model = lm(Petal.Length~Petal.Width,data = iris)
summary(linear_model) 
cor(iris$Petal.Length, iris$Petal.Width) #correlation coefficient (r)
plot(iris$Petal.Length, linear_model$residuals, xlab = "Petal length", ylab = "Residuals") #plot residuals against y
abline(h = 0, col = "darkgreen")
```

#### 6.1 模拟丢硬币
```{r}
set.seed(1) #设置随机数种子，使结果可复现 (reproducible research)
sample(c(0, 1), 1, replace = T, prob = c(0.2, 0.8)) #丢一枚作弊后的硬币
sample1 = sample(c(0, 1), 10, replace = T) #丢10枚正常硬币
table(sample1) / 10 #用一个表格总结概率
```

#### 6.2 概率
```{r}
#Suppose we have 10 trials with P(event) = 0.3
dbinom(4, 10, 0.3) #event发生4次的概率
pbinom(5, 10, 0.3) #event发生<=5次的概率
```

#### 7.1 高级一点的模拟丢硬币
```{r}
box = c(0,1) #建一个box model（假设丢一枚正常硬币），或者任何binary event
n = 10 #sample size是10
mean_box = mean(box)
mean_box
library(rafalib)
sd_box = popsd(box)
sd_box

#Sample sum:
n * mean_box #EV
sd_box * sqrt(n) #SE

#Sample mean:
mean_box #EV
sd_box/sqrt(n) #SE

totals = replicate(10000, sum(sample(box, 50, replace = T))) #丢50个硬币，重复10000次
table(totals) #用一个表格总结次数
hist(totals) #画一个histogram
```

#### 7.2 模拟1和0
```{r}
box = c(rep(1, 200), rep(0, 250)) #200个1，250个0
n = 30 #随机抽30个人 (sample size)
mean_box = mean(box)
sd_box = sd(box)

ev = mean_box 
ev #是1的概率
se = sd_box / sqrt(n) 
se #standard error

c(ev - 1*se, ev + 1*se) #68% confidence interval
c(ev - 2*se, ev + 2*se) #95% confidence interval
c(ev - 3*se, ev + 3*se) #99.7% confidence interval
```

#### 9.1 Hypothesis testing
```{r}
#H0: P = 0.8
#H1: P > 0.8
box = c(rep(0, 2), rep(1, 8))
mean(box)
library(rafalib)
popsd(box)
#sample size 29, 26个1
n = 29
ov = 26
ev = n * mean(box)
se = sqrt(n) * popsd(box)
ts = (ov-ev) / se #test statistic
pnorm(ts, lower.tail = FALSE) #P(test statistic >= 1.3)
#p-value > 0.05, retain H0
```

#### 10.1 1 Sample Z test
```{r}
#caffeine exhaustion time
caf0 = c(36.05,52.47,56.55,45.2,35.25,66.38,40.57,57.15,28.34)
caf13 = c(37.55,59.3,79.12,58.33,70.54,69.47,46.48,66.35,36.2)
#H0: mean exhaustion time with no caffeine = 45 (mins)
#H1: mean exhaustion time with no caffeine != 45 (mins)
teststat = (mean(caf0)-45)/(12/sqrt(length(caf0)))
teststat
2*pnorm(teststat, lower.tail = F)
#0.72 > 0.05, data appears consistent with the null hypothesis
```

#### 10.2 1 Sample T test
```{r}
teststat1 = (mean(caf0)-45) / (sd(caf0)/sqrt(length(caf0)))
teststat1
pt(teststat1, 8, lower.tail=F) #1 tailed, 8 = n-1
2*pt(teststat1, 8, lower.tail=F) #2 tailed
#p value for t test is slightly larger than z test, due to the extra wideness for a small sample size of n = 9 making the tail area larger
#conclusion is the same: we retain H0

#speedy way
t.test(caf0, mu = 45)
#shows: t = 0.34592, p-value = 0.7383
#95 percent confidence interval is same as 2 sided test with a = 0.05. As 45 falls inside the CI, we retain H0. 
```

#### 10.3 Paired T test
```{r}
cafdiff = caf13-caf0 #This is 1 sample! 
t.test(cafdiff, mu = 0)
#As the p-value is very small (0.01166), we reject H0 and conclude that "there is a difference". 
```

#### 10.4 2 Sample T test
```{r}
#Heart rates (beats per minute) for 2 independent groups of Sydney students, collected 20 minutes after the ‘RB’ group had drunk a 250ml cold can of Red Bull.
No_RB <- c(84,76,68,80,64,62,74,84,68,96,80,64,65,66)
RB <- c(72,88,72,88,76,75,84,80,60,96,80,84)

RB_data <- data.frame(group = c(rep("No_RB",14), rep("RB",12)), rate = c(No_RB,RB))

#H0: there is no difference between the mean heart rate of RB and No_RB
#H1: there is difference

#By formula
m1 = mean(No_RB)
m2 = mean(RB)
n1 = length(No_RB)
n2 = length(RB)
sdp2 = ((n1-1) * sd(No_RB)^2 + (n2-1) * sd(RB)^2)/(n1+n2-2)
se = sqrt(sdp2 * (1/n1 + 1/n2))
TestStat=(m1 - m2 - 0)/se
Pvalue=2*pt(abs(TestStat),n1+n2-2,lower.tail=F)
c(TestStat,Pvalue)

#Using shortcut
t.test(No_RB, RB, var.equal = T) #var.equal = T indicates that we are doing a two sample T test assuming equal population variances.

#Statistical conclusion: p-value > 0.05, retain H0. Data is consistent with hypothesis that mean heart rates are equal.
#Scientific conclusion: The data suggests that the consumption of Redbull by university students does not have an effect on heart rate. 
```

#### 10.5 Checking assumptions
```{r}
#Comparative boxplots - for normality and equality of variance assumptions
ggplot(RB_data, aes(x = group, y = rate, fill = group)) + geom_boxplot() + 
  ggtitle("Boxplot of effect of Red Bull on student heart rate") + labs(x = "Group", y = "Heart rate (bpm)") + 
  theme(legend.position = "none")

#QQ-plot - for normality
ggplot(RB_data, aes(sample = rate, colour = group)) + stat_qq() + stat_qq_line() + ggtitle("QQplot")

#Shapiro-Wilk test - for normality
shapiro.test(No_RB)
shapiro.test(RB)

#Levene's test (F-test) - for equal spread
var.test(No_RB,RB)
#The F-test gives p-value > 0.05, suggesting that data is consistent with equal variance.
#If variance is different, we ma yincrease the chance of falsely rejecting H0.

#Suppose variance is different. 
set.seed(10)
No_RB <- No_RB + rnorm(14, 0, 20)  #New simulated version of No-RB with larger variance
var.test(No_RB, RB)
t.test(No_RB, RB, var.equal = FALSE) #This performs the Welch 2 Sample T Test. 
# p-value > 0.05, so our conclusion concurs with the previous 2 Sample T Test for equal variances. 
```
#### 11.1 Chi-square test - Goodness of Fit
```{r}
set.seed(1)
dice=c(1,2,3,4,5,6)
fair=c(1/6,1/6,1/6,1/6,1/6,1/6)
throws1=sample(dice, 60,replace = T,prob=fair)
throws1a = table(throws1)
throws1a
sum(((table(throws1)-10)^2)/10)
pchisq(4.4,5,lower.tail=F)

#Speedy way
chisq.test(throws1a, p = c(1/6,1/6,1/6,1/6,1/6,1/6))

#Loaded dice
set.seed(1)
dice=c(1,2,3,4,5,6)
loaded=c(1/6-1/7,1/6,1/6,1/6,1/6,1/6+1/7)
throws2=sample(dice, 60,replace = T,prob=loaded)
throws2a <- table(throws2)
throws2a
chisq.test(throws2a, p = c(1/6,1/6,1/6,1/6,1/6,1/6))
```
#### 11.2 Chi-square test - Independence
```{r}
Smoke = matrix(c(9,4,6,6), nrow = 2, ncol = 2, byrow = TRUE, dimnames = list(c("Smoke", "No smoke"), c("Domestic", "International")))
mosaicplot(Smoke)
chisq.test(Smoke)
chisq.test(Smoke, correct = F) # Without continuity correction
mosaicplot(Smoke, shade = T) # Mosaic plot with standardised residuals
fisher.test(Smoke) #If you have a small sample size (cells < 5), and you cannot increase the sample size by, for example, combining categories, then you can use Fisher’s exact test fishers.test in R. 
```

#### 11.3 Regression Test

```{r}
lm(Petal.Length~Petal.Width, iris)
summary(lm(Petal.Length~Petal.Width, iris))
```

